Metabolites reflect variability introduced by mesoscale eddies in the North Pacific Subtropical Gyre

### 2024-11-15 manuscript/response_to_reviewers.txt

> Reviewer 1 | 04 Sep 2024 | 01:18
> #1
> This study examines the impact of mesoscale eddies on the metabolome and composition 
> of particulate organic matter in the North Pacific Subtropical Gyre. It fills a gap 
> in previous research by exploring the molecular composition of particulate organic 
> matter. The use both of targeted and untargeted metabolomics mass spectrometry to 
> detect molecules sensitive to environmental perturbation. The study finds that many 
> metabolites track biomass trends. Metabolites linked to eukaryotic phytoplankton are 
> enriched at the deep chlorophyll maximum of cyclonic features, while degradation 
> products are enriched in anticyclones. Additionally, depth variability and vertical 
> displacement of isopycnal layer constrain the importance of mesoscale eddies in 
> shaping particulate matter composition. Based on these results, we can infer the 
> biological communities and biogeochemical processes driven by mesoscale eddies at 
> both spatial and temporal scales. Viewing this complex dynamic process from the 
> chemical molecular scale is a perspective that sets this article apart from other 
> traditional methods in examining the changes in this system.
This is an accurate assessment of the manuscript.

> However, I still have some questions.
> Currently, the process of the entire system seems to be predictable. Although the 
> author has introduced and discussed some gaps, especially from the perspective of 
> chemical composition, and this article provides detailed data and reliable analysis, 
> it is not prominent enough from the perspective of the entire system. Perhaps the 
> transformation from space to time and the use of chemical composition to observe the 
> aging process of vortices is a great highlight. It is very difficult to focus on 
> scale eddies and the ecosystem evolution driven by them using traditional methods. 
> If that's the case, the structure and writing style of the paper may need to be 
> adjusted.
> This article focuses on the differences in chemical composition to understand this 
> process, but I always want to connect it with specific biological processes, such 
> as structural changes in biological communities, primary production rates, 
> zooplankton grazing rates, or bacterial production and respiration rates etc. 
> The author also mentioned these. I don't know if the author can add more relevant 
> evidence or discussion to allow readers to connect with these specific biological 
> groups and/or processes. I know it's difficult to integrate these data together, 
> but relevant discussions are possible.
> We already know that mesoscale eddies are complex and dynamic multidimensional 
> processes. The three-dimensional space combined with the temporal evolution of 
> ecosystem structure and function. The changes in the density surface driven by 
> physics during this process lead to the coupling reaction of physics chemistry 
> biology, which is the main thread. In fact, many other physical biological coupling 
> processes are similar, such as upwelling, internal waves/tides. Using the author's 
> method, I think similar situations can also be seen in such processes.
We have expanded the Discussion to touch on the concerns raised by this reviewer,
adding section 5.3 to respond to this directly.
Specifically, we have mentioned possible links between the increased sulfonium
compounds in the eddy and DMS production, a hypothesis about the reason why
increased cyclonic biomass was not matched by increased export, and extensive
references to additional manuscripts with relevant data in lines 651-655 and 
689-718.

@@

> Blanks are mentioned but it is not shown how these were used for the data analysis.
We have added a note about how we used the methodological blanks to visually assess 
whether sample peak areas were above the background signal. (Lines 184-186)

> In line 142 the method for the evaluation of the matrix effect is mentioned, but no 
> results are given.
The Boysen et al. (2018) paper details the method for identifying the magnitude of the 
matrix effect by using a full-strength and water-diluted pooled sample and we believe 
that the reader is best suited to the detailed analysis presented there than a
brief review of it in this manuscript. We have confirmed that references to
this method are cited immediately after its mention here.

> The experimental could be clearer in line 132 ff “The samples were then bead-beaten 
> three times, followed by triplicate extraction of the aqueous layer...” Was the 
> organic layer collected or was the aqueous layer collected (as suggested by the rest 
> of the sentence?). Was phase separation occurring spontaneously or was a 
> centrifugation required? Was separation done by pipetting?
We have added additional text to this line to clarify the requested points of 
confusion. (Lines 133-138)

> Line 177: I would rather state “targeted analytical methods”. Metabolomics is by 
> definition rather untargeted.
We have restructured the headings for this section of the methods and added
additional detail to respond to this concern.

### 2024-11-15 manuscript/manuscript.Rmd
Switching section headings per Rachel and reviewer recs
Adding Anitra and Benedetto recs for section 5.3

## Targeted metabolomic methods

@@

## Untargeted metabolomic methods

@@

Missing links between the environment with its associated community and the composition of particulate matter make it difficult to estimate important controls on the marine carbon cycle. This is especially true for dynamic and short-lived environments such as eddies. Metabolites offer a way to bridge this gap in knowledge by directly measuring the units of exchange between organisms and their environment.

Here, we saw how changes in the density surface driven by physics during this process led to the coupling of physics, chemistry, and biology. The enrichment of eukaryotic organisms at the DCM of the cyclone mirrors other ocean processes where upwelling of deep, nutrient-rich water leads to increases in nitrogen-rich metabolites representative of a shift towards organisms such as diatoms and dinoflagellates that undergo bloom dynamics [@Durham2022; @Kuhlisch2023]. @Barone2022 established a "chemical wake" hypothesis in which cyclone formation introduces a transient phytoplankton community with net photosynthetic production that induces export, a conceptual model that can be extended to include many of the metabolites named here, as some metabolites are more bioavailable than others and the "flavor" of this chemical wake will change over time.

A shift from targeted to untargeted metabolites may also in part explain why the increased biomass in the cyclone did not result in large increases in export [@Barone2022; @Bent2024]. Our list of known compounds covers a large fraction of the small molecules common to most organisms, indicating that unknown molecules are more likely to be rare and therefore more difficult to degrade or a result of prokaryote dominance. If the increased biomass in the cyclone is dominantly in the form of highly bioavailable compounds, it is not surprising that much of it can be easily degraded. This hypothesis is supported by @Gleich2024, who found that heterotrophy-associated transcripts were more abundant in the 2018 cruise cyclone than the anticyclone.

Almost all large phytoplankton classes in the 2017 eddy pair were enriched in the cyclone DCM relative to the anticyclone (particularly prymnesiophytes, dinoflagellates, diatoms, and green algae), while *Prochlorococcus* was significantly less abundant and *Synechococcus* was approximately equal [@Barone2022]. Thus, the metabolite data presented here does not help distinguish between large phytoplankton classes but instead provides additional cues about the shift from a metabolome dominated by cyanobacteria producers into one dominated by eukaryotic organisms. This shift occurs naturally over the life cycle of an eddy, with distinctions typically drawn between the eddy intensification, its mature phase, and eventual decay [@Rii2008; @McGillicuddy2016; @Kuhlisch2023; @Barone2022]. The initial introduction of inorganic nutrients during intensification of cyclonic eddies followed by diapycnal mixing can still stimulate the DCM community without large effects on integrate primary productivity or export even months after formation, as seen here. Both cruises occurred in weakening cyclones and mature anticyclones, limiting our ability to make direct conclusions about all eddy life cycle phases. However, the general alignment of the results presented here with our initial hypotheses strengthens our confidence in metabolites as links between eddy biogeochemistry and the associated biological shifts anywhere.

>>>>

## Metabolomic data analysis

### Full scan feature extraction for known molecules

@@

### Automated processing for detection of unknown features

@@

The composition of particulate matter controls the turnover of organic carbon in the ocean, but it is still unclear how this composition is altered in the environment and the plankton community. Measuring metabolites can quantify the variability of the central molecules responsible for the flow of energy and matter through the marine ecosystem at many different scales. In these results, we focused on mesoscale eddies, which perturb the ocean ecosystem on relatively small horizontal scales on the order of tens to hundreds of kilometers.

Here, we reported how changes in the distribution of water layers of different density driven by mesoscale physical dynamics cascading effects in the biological and chemical characteristics of the eddies. The enrichment of eukaryotic organisms at the DCM of the cyclone mirrors other ocean processes where upwelling of deep, nutrient-rich water leads to increases in sulfur-rich metabolites representative of a shift towards organisms such as diatoms and dinoflagellates that undergo bloom dynamics [@Durham2022; @Kuhlisch2023]. @Barone2022 proposed that a "chemical wake" is formed after cyclone intensification when a transient phytoplankton community with net photosynthetic production induces export and leaves deficits and excesses of inorganic nutrients and oxygen, respectively. This conceptual model that can be extended to include many of the metabolites named here, as some metabolites are more bioavailable than others and the "flavor" of the chemical wake will change throughout the eddy life cycle.

The prevalence of unknown metabolites relative to known metabolites in the anticyclone may also in part explain why the increased biomass in the cyclone did not result in large increases in export [@Barone2022; @Bent2024]. Our list of known compounds covers a large fraction of the small molecules common to most organisms, indicating that unknown molecules are more likely to be rare and therefore more difficult to degrade or a result of prokaryote dominance. If the increased biomass in the cyclone is dominantly in the form of highly bioavailable compounds, it is not surprising that much of it can be easily degraded. This hypothesis is supported by @Gleich2024, who found that heterotrophy-associated transcripts were more abundant in the 2018 cruise cyclone than the anticyclone, indicating that there was fresh material for the heterotrophic bacteria to readily consume.

Almost all large phytoplankton classes in the 2017 eddy pair were enriched in the cyclone DCM relative to the anticyclone (particularly prymnesiophytes, dinoflagellates, diatoms, and green algae), while *Prochlorococcus* was significantly less abundant and *Synechococcus* was approximately equal [@Barone2022]. Thus, the metabolite data presented here does not help distinguish between large phytoplankton classes but instead provides additional cues about the shift from a metabolome dominated by cyanobacteria producers into one dominated by eukaryotic organisms. This shift occurs naturally over the life cycle of an eddy, with distinctions typically drawn between the eddy intensification, its mature phase, and eventual decay [@Rii2008; @McGillicuddy2016; @Kuhlisch2023; @Barone2022]. Increased diapycnal mixing of nutrients in mature cyclones can alter the DCM community without large effects on integrated primary productivity or export even months after eddy intensification [@Barone2022]. We here demonstrated how this dynamic is linked with large changes in the chemical characteristics of the particulate material.
